---
title: "The First Project"
author: "Ekaterina Kravchuk"
date: "05 11 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Olimpic games

Working with data on Olympic Games participants over the past 120 years

## 0. Installing packages

```{r}
libraries <- c("skimr", "dplyr")
loading_libraries <- function(libraries) {
  installed_libraries <- libraries %in% rownames(installed.packages())
  if (any(installed_libraries == FALSE)) {
  install.packages(libraries[!installed_libraries])
    }
  invisible(lapply(libraries, library, character.only = TRUE))
}
loading_libraries(libraries)
```

## 1. Getting the data

A function for combining observations into one table

```{r, echo=TRUE}
getting_data <- function(path, type ) {
  all_files <- list.files(path, pattern = type, full.names = TRUE)
  listed_files <- lapply(all_files, read.csv)
  dataset <- lapply(listed_files, rbind)  
  do.call(rbind.data.frame, dataset)
}

olympic_data_table <- getting_data('~/data', '.csv')
```
## 2. Data correction

```{r, echo=TRUE}
str(olympic_data_table)
```
We have 271115 obs. of  15 variables
```{r, echo = TRUE}
library(skimr)
skim(olympic_data_table)

```
Let's replace empty strings with NA:

```{r}
olympic_tidy_data <- olympic_data_table %>% replace(. == "", NA)
```
Let's remove the column 'Games', we have the same information in 'Year' and 'Season'

```{r}
olympic_tidy_data$Games = NULL
```
Let's make 'Sex' and 'Medal', 'Season' factors
```{r, echo = TRUE}
unique(olympic_tidy_data$Sex)
```
What does 'G' mean?
```{r, echo = TRUE}
subset(olympic_tidy_data, olympic_tidy_data$Sex == "G" | is.na(olympic_tidy_data$Sex))
```
We see male names. Let's replace 'G' with 'M'
```{r}
olympic_tidy_data$Sex = replace(olympic_tidy_data$Sex, olympic_tidy_data$Sex == "G", "M")
olympic_tidy_data$Sex = as.factor(olympic_tidy_data$Sex)
```
```{r, echo = TRUE}
unique(olympic_tidy_data$Season)
```
```{r}
olympic_tidy_data$Season = as.factor(olympic_tidy_data$Season)
```
```{r, echo = TRUE}
unique(olympic_tidy_data$Medal)
```
```{r}
olympic_tidy_data$Medal = as.factor(olympic_tidy_data$Medal)
```
Let's change 'Height' to numeric
```{r}
olympic_tidy_data$Height = as.numeric(olympic_tidy_data$Height)
```
Let's look at the summary of the data we have now
```{r, echo = TRUE}
summary(olympic_tidy_data)
```
I saved this data
```{r}
write.csv(olympic_tidy_data, file = "~/data/olympic_tidy_data.csv")
```
## 3. Finding out the age of the youngest athletes of both sexes at the 1992 Olympics
```{r, echo = TRUE}
library(dplyr)
young = olympic_tidy_data %>% 
  filter(Year == 1992) %>% 
  group_by(Sex) %>% 
  slice_min(order_by = Age)
select(young, Name, Sex, Age)
```
We see, the youngest male is 11, the youngest female is 12

## 4. Calculating the mean and standard deviation of the Height variable for athletes of each gender
```{r, echo = TRUE}
olympic_tidy_data %>% 
  group_by(Sex) %>% 
  summarise(mean = mean(Height, na.rm = TRUE), sd = sd(Height, na.rm = TRUE))
```
## 5. Calculating the average value and standard deviation of the Height variable for female tennis players (sex = F) at the 2000 Olympics
```{r, echo = TRUE}
olympic_tidy_data %>% 
  filter(Sex == "F" & Sport == "Tennis" & Year == 2000) %>%
    summarise(mean = round(mean(Height, na.rm = TRUE), 1), sd = round(sd(Height, na.rm = TRUE), 1))
```
## 6. What kind of sport did the heaviest athlete participate in at the 2006 Olympics?
```{r, echo = TRUE}
the_haviest = olympic_tidy_data %>% 
  filter(Year == 2006) %>% 
  slice_max(order_by = Weight)
the_haviest$Sport
```
## 7. How many gold medals were received by women from 1980 to 2010?
```{r, echo = TRUE}
olympic_tidy_data %>% 
  filter(Year %in% c(1980:2010) & Sex == 'F' & Medal == "Gold") %>%
  nrow()
```
## 8. How many times did John Aalberg participate in Olympiad?
```{r, echo = TRUE}
olympic_tidy_data %>% 
  filter(Name == "John Aalberg") %>%
  nrow()
```